$Z^\prime$ mass limits and the naturalness of supersymmetry

### 2015-03-27 paper/e6ssm-tuning-paper.tex
Fixed type in one-loop beta for m_S^2, restored two-loop trace terms to two-loop betas for \lambda and y_t that are missing in the trilinears
Removed brackets from SU(2) dot products but left the dots, added expression for loop contributions to effective potential, changed reference to NMSSM tuning paper to indicate that it has been previously used

Here we will refer to U(1) extensions of the MSSM, which solve the
$\mu$-problem, as the USSM \cite{Cvetic:1995rj, Jain:1995cb,
  Nir:1995bu, Cvetic:1996mf, Cvetic:1997ky}.  The couplings for the
$U(1)^\prime$ gauge group should allow the following renormalisable
superpotential terms required in the USSM, \be W_{USSM} = y^U_{ij}
\hat{u}^c_i \left ( \hat{H}_u \cdot \hat{Q}_j \right ) + y^D_{ij} \hat{d}^c_i 
\left ( \hat{Q}_j \cdot \hat{H}_d \right )
 + y^E_{ij} \hat{e}^c_i \left ( \hat{L}_j \cdot \hat{H}_d \right ) + \lambda \hat{S}
\left ( \hat{H}_d \cdot \hat{H}_u \right ), \ee with $i,j \in \{1,2,3\}$. 
For the $SU(2)$ dot product we follow the convention $\hat{A}\cdot \hat{B}\equiv \epsilon_{\alpha\beta}\hat{A}^\alpha\hat{B}^\beta=\hat{A}^2\hat{B}^1-\hat{A}^1\hat{B}^2$.

@@

The breaking of $E_6$ into $SO(10)$, gives rise to $E_6\to
SO(10)\times U(1)_{\psi}$, and the subsequent breaking of $SO(10)$
into $SU(5)$, gives $SO(10)\to SU(5)\times U(1)_{\chi}$ (this is
reviewed in e.g.~\cite{Langacker:2008yv}). The extra $U(1)$ gauge
symmetry at low energies should then be a linear combination of these
in the $E_6$ inspired case, \be U(1)^\prime = \cos(\theta) U(1)_{\chi}
+ \sin(\theta) U(1)_{\psi}. \ee In Table \ref{tab:E6charges} the
charges for several popular $E_6$ inspired $U(1)$ extensions are
shown.
% Table of U(1)' charges, which may or may not be useful to include.
\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cccccccccccccc}
 & $\hat{Q}$ & $\hat{u}^C$ & $\hat{d}^C$ & $\hat{L}$ & $\hat{e}^C$ & $\hat{N}^C$ & $\hat{S}$ & $\hat{H}_2$ & $\hat{H}_1$ & $\hat{D}$ & $\hat{\overline{D}}$ & $\hat{H}'$ & $\hat{\overline{H'}}$ \\[1mm]
\hline
$\sqrt{\frac{5}{3}}Q_i^Y$ & $\frac{1}{6}$ & $-\frac{2}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $1$ & $0$ & $0$ & $\frac{1}{2}$ & $-\frac{1}{2}$ & $-\frac{1}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $\frac{1}{2}$ \\[1mm]
%\hline
$2\sqrt{6}Q_i^\psi$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $4$ & $-2$ & $-2$ & $-2$ & $-2$ & $1$ & $-1$\\[1mm]
%\hline
$2\sqrt{10}Q_i^\chi$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ & $-5$ & $0$ & $2$ & $-2$ & $2$ & $-2$ & $3$ & $-3$\\[1mm]
%\hline
$\sqrt{40}Q_i^N$ & $1$ & $1$ & $2$ & $2$ & $1$ & $0$ & $5$ & $-2$ & $-3$ & $-2$ & $-3$ & $2$ & $-2$ \\[1mm]
\end{tabular}
\end{ruledtabular}
\caption{The $U(1)_Y$, $U(1)_\psi$, $U(1)_\chi$ and $U(1)_N$ charges of the chiral superfields in the $E_6$ model. The specific case of $U(1)_N$, corresponding to the E$_6$SSM, is obtained for $\theta=\arctan\sqrt{15}$.}
\label{tab:E6charges}
\end{table}

@@

\be
W_{E_6} = W_0 + W_1 + W_2,
\label{Eq:e6-superpot}
\ee where
\begin{eqnarray}
W_0 &=& \lambda_{ijk} \hat{S}_i \left ( \hat{H}^d_{j} \cdot\hat{H}^u_{k}\right ) + \kappa_{ijk} \hat{S}_i \hat{D}_j \hat{\bar{D}}_k + h^N_{ijk} \hat{N}^c_i \left ( \hat{H}^u_{j} \cdot \hat{L}_k\right ) \nonumber \\
& & + h^U_{ijk} \hat{u}^c_i \left ( \hat{H}^u_{j} \cdot \hat{Q}_k\right ) + h^D_{ijk} \hat{d}^c_i \left ( \hat{Q}_k\cdot\hat{H}^d_{j}\right )  + h^E_{ijk} \hat{e}^c_i \left ( \hat{L}_k\cdot\hat{H}^d_{j}\right ) , \\
W_1 &=& g^Q_{ijk} \hat{D}_i \left ( \hat{Q}_j \cdot \hat{Q}_k \right )+ g^q_{ijk} \hat{\bar{D}}_i \hat{d}^c_j \hat{u}^c_k, \\
W_2 &=& g^N_{ijk} \hat{N}^c_i \hat{D}_j \hat{d}^c_k + g^E_{ijk} \hat{e}^c_i \hat{D}_j \hat{u}^c_k + g^D_{ijk} \left ( \hat{Q}_i \cdot \hat{L}_j \right )\hat{\bar{D}}_k.
\label{Eq:e6-superpot-parts}
\end{eqnarray}

Nonetheless while this model is very elegant so far, the
superpotential of Eq.~(\ref{Eq:e6-superpot}) contains dangerous terms which
can induce proton decay and lead to large flavour changing neutral
currents (FCNCs).  There are a number of approaches to suppress these
terms, involving the use of different discrete symmetries.  Here for
the purposes of renormalisation group running we will simply include
the following unsuppressed superpotential terms, which follows the
approach taken in work on the cE$_6$SSM \cite{Athron:2009ue,
  Athron:2009bs},
%
%% \begin{align}
%%   \begin{split}
    \ba
    W &\approx& y_{\tau} \left (\SuperField{L}_3 \cdot \SuperField{H}_d
     \right ) \SuperField{e}^C_3 + y_b \left (\SuperField{Q}_3\cdot\SuperField{H}_d
     \right ) \SuperField{d}_3^C + y_t \left (\SuperField{H}_u\cdot\SuperField{Q}_3
     \right ) \SuperField{u}_3^C\\
    &&
    + \lambda_i \SuperField{S} \left (\SuperField{H}_{1i}
    \cdot \SuperField{H}_{2i} \right ) + \kappa_i \SuperField{S} \SuperField{D}_i
    \SuperField{\overline{D}}_i + \mu' \left (\SuperField{H}'\cdot
    \SuperField{\overline{H'}}\right ),
    \ea
  %% \end{split}
  \label{SuPot_RGE}
%%\end{align}
%

@@

V_F&=\lambda^2|S|^2(|H_1|^2+|H_2|^2)+\lambda^2|(H_1\cdot
H_2)|^2,\label{eq:E6VFterms}\\ V_D&=\frac{\bar{g}^2}{8}\left (

@@

H_2|^2+\frac{g_1'^2}{2}(Q_1|H_1|^2+Q_2|H_2|^2+Q_S|S|^2)^2,\label{eq:E6VDterms}\\ V_{\textrm{soft}}&=m_S^2|S|^2+m_{H_d}^2|H_1|^2+m_{H_u}^2|H_2|^2+\Big
[\lambda A_\lambda S(H_1\cdot H_2)+\textrm{h.c.}\Big

@@

$H_1$, $H_2$ and $S$, respectively. The term $\Delta V$ contains the
Coleman-Weinberg contributions to the effective potential. In the case

@@

$\mu$ term, as occurs in the MSSM.\\ \\ Demanding that
the Higgs fields $H_1$, $H_2$ and the singlet $S$ acquire real VEVs of
the form

@@

(v_2^2-v_1^2)v_1+D_{H_1}v_1+\frac{\partial \Delta V_{E_6}}{\partial v_1}=0,\label{eq:E6EWSBcondition1} \\

@@

(v_2^2-v_1^2)v_2+D_{H_2}v_2+\frac{\partial \Delta V_{E_6}}{\partial v_2}=0,\label{eq:E6EWSBcondition2} \\
&f_3=m_S^2s+\frac{\lambda^2}{2}(v_2^2+v_1^2)s-\frac{\lambda A_\lambda}{\sqrt{2}}v_2v_1+D_Ss+\frac{\partial \Delta V_{E_6}}{\partial s}=0.\label{eq:E6EWSBcondition3}

@@

D_\phi\equiv \frac{g_1'^2}{2}\left (
Q_1v_1^2+Q_2v_2^2+Q_Ss^2\right
)Q_\phi.
As was noted in \cite{Athron:2013ipa}, the first two of these
conditions may be rewritten in the form

@@

Eq. (\ref{eq:E6MZequation}). The general procedure is as follows
\cite{Ellwanger:2011mu}. For a model in which $m$ fields develop real

@@

&{}+\frac{9}{5}g_1^2g_2^2+6g_1'^2g_2^2\left ( Q_1^2+Q_2^2\right )+\frac{6}{5}g_1^2g_1'^2\left (Q_1^2+Q_2^2\right )\bigg \},\label{eq:USSMLambdaBetaTwoLoop}\\

@@

\beta_{m_S^2}^{(1)}&=\sum_{i=1}^3\left [ 4\lambda_i^2\left ( m_{H_i^d}^2+m_{H_i^u}^2+m_S^2\right )+4a_{\lambda_i}^2+6\kappa_i\left ( m_S^2+m_{D_i}^2+m_{\overline{D}_i}^2\right )+6a_{\kappa_i}^2\right ]\nonumber\\

@@

&{}+\frac{2}{5}g_1^2g_1'^2\left [ 3Q_2^2+\frac{1}{3}Q_Q^2+\frac{16}{3}Q_u^2\right ]\bigg \},\label{eq:USSMYu22BetaTwoLoop}\\

>>>>

Here we will refer to U(1) extensions of the MSSM, which solve the
$\mu$-problem, as the USSM \cite{Cvetic:1995rj, Jain:1995cb,
  Nir:1995bu, Cvetic:1996mf, Cvetic:1997ky}.  The couplings for the
$U(1)^\prime$ gauge group should allow the following renormalisable
superpotential terms required in the USSM, \begin{equation}
 W_{USSM} = y^U_{ij}
\hat{u}^c_i  \hat{H}_u \cdot \hat{Q}_j + y^D_{ij} \hat{d}^c_i  \hat{Q}_j \cdot \hat{H}_d
 + y^E_{ij} \hat{e}^c_i \hat{L}_j \cdot \hat{H}_d  + \lambda \hat{S} \hat{H}_d \cdot \hat{H}_u, 
\end{equation} with $i,j \in \{1,2,3\}$. 
For the $SU(2)$ dot product we follow the convention $\hat{A}\cdot \hat{B}\equiv \epsilon_{\alpha\beta}\hat{A}^\alpha\hat{B}^\beta=\hat{A}^2\hat{B}^1-\hat{A}^1\hat{B}^2$.

@@

The breaking of $E_6$ into $SO(10)$, gives rise to $E_6\to
SO(10)\times U(1)_{\psi}$, and the subsequent breaking of $SO(10)$
into $SU(5)$, gives $SO(10)\to SU(5)\times U(1)_{\chi}$ (this is
reviewed in e.g.~\cite{Langacker:2008yv}). The extra $U(1)$ gauge
symmetry at low energies should then be a linear combination of these
in the $E_6$ inspired case, \be U(1)^\prime = U(1)_{\chi}\cos\theta 
+ U(1)_{\psi}\sin\theta . \ee In Table \ref{tab:E6charges} the
charges for several popular $E_6$ inspired $U(1)$ extensions are
shown.
% Table of U(1)' charges, which may or may not be useful to include.
\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cccccccccccccc}
 & $\hat{Q}$ & $\hat{u}^C$ & $\hat{d}^C$ & $\hat{L}$ & $\hat{e}^C$ & $\hat{N}^C$ & $\hat{S}$ & $\hat{H}_2$ & $\hat{H}_1$ & $\hat{D}$ & $\hat{\overline{D}}$ & $\hat{H}'$ & $\hat{\overline{H'}}$ \\[1mm]
\hline
$\sqrt{\frac{5}{3}}Q_i^Y$ & $\frac{1}{6}$ & $-\frac{2}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $1$ & $0$ & $0$ & $\frac{1}{2}$ & $-\frac{1}{2}$ & $-\frac{1}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $\frac{1}{2}$ \\[1mm]
%\hline
$2\sqrt{6}Q_i^\psi$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $4$ & $-2$ & $-2$ & $-2$ & $-2$ & $1$ & $-1$\\[1mm]
%\hline
$2\sqrt{10}Q_i^\chi$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ & $-5$ & $0$ & $2$ & $-2$ & $2$ & $-2$ & $3$ & $-3$\\[1mm]
%\hline
$\sqrt{40}Q_i^N$ & $1$ & $1$ & $2$ & $2$ & $1$ & $0$ & $5$ & $-2$ & $-3$ & $-2$ & $-3$ & $2$ & $-2$ \\[1mm]
\end{tabular}
\end{ruledtabular}
\caption{The $U(1)_Y$, $U(1)_\psi$, $U(1)_\chi$ and $U(1)_N$ charges of the chiral superfields in the $E_6$ model. The specific case of $U(1)_N$, corresponding to the E$_6$SSM, is obtained for $\theta=\arctan\sqrt{15}$.}
\label{tab:E6charges}
\end{table}

@@

\be
W_{E_6} = W_0 + W_1 + W_2,
\label{Eq:e6-superpot}
\ee where
\begin{eqnarray}
W_0 &=& \lambda_{ijk} \hat{S}_i  \hat{H}^d_{j} \cdot\hat{H}^u_{k} + \kappa_{ijk} \hat{S}_i \hat{D}_j \hat{\bar{D}}_k + h^N_{ijk} \hat{N}^c_i  \hat{H}^u_{j} \cdot \hat{L}_k \nonumber \\
& & + h^U_{ijk} \hat{u}^c_i  \hat{H}^u_{j} \cdot \hat{Q}_k + h^D_{ijk} \hat{d}^c_i  \hat{Q}_k\cdot\hat{H}^d_{j}  + h^E_{ijk} \hat{e}^c_i  \hat{L}_k\cdot\hat{H}^d_{j} , \\
W_1 &=& g^Q_{ijk} \hat{D}_i  \hat{Q}_j \cdot \hat{Q}_k + g^q_{ijk} \hat{\bar{D}}_i \hat{d}^c_j \hat{u}^c_k, \\
W_2 &=& g^N_{ijk} \hat{N}^c_i \hat{D}_j \hat{d}^c_k + g^E_{ijk} \hat{e}^c_i \hat{D}_j \hat{u}^c_k + g^D_{ijk}  \hat{Q}_i \cdot \hat{L}_j \hat{\bar{D}}_k.
\label{Eq:e6-superpot-parts}
\end{eqnarray}

Nonetheless while this model is very elegant so far, the
superpotential of Eq.~(\ref{Eq:e6-superpot}) contains dangerous terms which
can induce proton decay and lead to large flavour changing neutral
currents (FCNCs).  There are a number of approaches to suppress these
terms, involving the use of different discrete symmetries.  Here for
the purposes of renormalisation group running we will simply include
the following unsuppressed superpotential terms, which follows the
approach taken in work on the cE$_6$SSM \cite{Athron:2009ue,
  Athron:2009bs},
%
%% \begin{align}
%%   \begin{split}
    \ba
    W &\approx& y_{\tau} \SuperField{L}_3 \cdot \SuperField{H}_d
      \SuperField{e}^C_3 + y_b \SuperField{Q}_3\cdot\SuperField{H}_d
      \SuperField{d}_3^C + y_t \SuperField{H}_u\cdot\SuperField{Q}_3
      \SuperField{u}_3^C\\
    &&
    + \lambda_i \SuperField{S} \SuperField{H}_{1i}
    \cdot \SuperField{H}_{2i}  + \kappa_i \SuperField{S} \SuperField{D}_i
    \SuperField{\overline{D}}_i + \mu' \SuperField{H}'\cdot
    \SuperField{\overline{H'}},
    \ea
  %% \end{split}
  \label{SuPot_RGE}
%%\end{align}
%

@@

V_F&=\lambda^2|S|^2(|H_1|^2+|H_2|^2)+\lambda^2|H_1\cdot
H_2|^2,\label{eq:E6VFterms}\\ 
V_D&=\frac{\bar{g}^2}{8}\left (

@@

H_2|^2+\frac{g_1'^2}{2}(Q_1|H_1|^2+Q_2|H_2|^2+Q_S|S|^2)^2,\label{eq:E6VDterms}\\ 
V_{\textrm{soft}}&=m_S^2|S|^2+m_{H_d}^2|H_1|^2+m_{H_u}^2|H_2|^2+\Big
[\lambda A_\lambda SH_1\cdot H_2+\textrm{h.c.}\Big

@@

$H_1$, $H_2$ and $S$, respectively. In the case

@@

$\mu$ term, as occurs in the MSSM. The term $\Delta V$ contains the
Coleman-Weinberg contributions to the effective potential. For the purposes of this study, 
we include in $\Delta V$ only the one-loop contributions from the top quark and stop squarks,
\Delta V=\frac{3}{32\pi^2}\left [ m_{\tilde{t}_1}^4\left ( \log\frac{m_{\tilde{t}_1}^2}{Q^2}-\frac{3}{2}\right )+m_{\tilde{t}_2}^4\left ( \log\frac{m_{\tilde{t}_2}^2}{Q^2}-\frac{3}{2}\right )-2m_t^4\left ( \log\frac{m_t^2}{Q^2}-\frac{3}{2}\right )\right ].
Explicit expressions for the running $\overline{\textrm{DR}}$ top mass $m_t$ and stop masses $m_{\tilde{t}_{1,2}}$ are given 
below.\\ \\
Demanding that the Higgs fields $H_1$, $H_2$ and the singlet $S$ acquire real VEVs of the form

@@

(v_2^2-v_1^2)v_1+D_{H_1}v_1+\frac{\partial \Delta V}{\partial v_1}=0,\label{eq:E6EWSBcondition1} \\

@@

(v_2^2-v_1^2)v_2+D_{H_2}v_2+\frac{\partial \Delta V}{\partial v_2}=0,\label{eq:E6EWSBcondition2} \\
&f_3=m_S^2s+\frac{\lambda^2}{2}(v_2^2+v_1^2)s-\frac{\lambda A_\lambda}{\sqrt{2}}v_2v_1+D_Ss+\frac{\partial \Delta V}{\partial s}=0.\label{eq:E6EWSBcondition3}

@@

D_\phi\equiv \frac{g_1'^2}{2}\left (Q_1v_1^2+Q_2v_2^2+Q_Ss^2\right )Q_\phi.
We also include these $U(1)'$ $D$-term contributions in the diagonalised stop masses,
  m_{\tilde{t}_{1,2}}^2&=\frac{1}{2}\bigg \{ m_{Q3}^2+m_{u3}^2+\frac{1}{2}M_Z^2\cos 2\beta+D_Q+D_u+2m_t^2\nonumber\\
&\quad{}\mp\sqrt{\left [ m_{Q3}^2+m_{u3}^2+\left ( \frac{1}{2}-\frac{4}{3}\sin^2\theta_W\right )M_Z^2\cos 2\beta + D_Q-D_u\right ]^2+4m_t^2X_t^2}\bigg \},
where $m_t^2=y_t^2v_2^2/2$ and $X_t=A_t-\frac{\lambda s v_1}{\sqrt{2}v_2}$. By definition we take $m_{\tilde{t}_1}$ to correspond to the lighter of the two states.\\ \\
As was noted in \cite{Athron:2013ipa}, the first two of these conditions may be rewritten in the form

@@

Eq. (\ref{eq:E6MZequation}). The general procedure that we use is as follows
(this method has also previously been applied in the NMSSM, see for example \cite{Ellwanger:2011mu}). For a model in which $m$ fields develop real

@@

&{}+\frac{9}{5}g_1^2g_2^2+6g_1'^2g_2^2\left ( Q_1^2+Q_2^2\right )+\frac{6}{5}g_1^2g_1'^2\left [Q_1^2+Q_2^2+\left ( Q_2 - Q_1 \right )\Sigma_Q^Y\right ]\bigg \},\label{eq:USSMLambdaBetaTwoLoop}\\

@@

\beta_{m_S^2}^{(1)}&=\sum_{i=1}^3\left [ 4\lambda_i^2\left ( m_{H_i^d}^2+m_{H_i^u}^2+m_S^2\right )+4a_{\lambda_i}^2+6\kappa_i^2\left ( m_S^2+m_{D_i}^2+m_{\overline{D}_i}^2\right )+6a_{\kappa_i}^2\right ]\nonumber\\

@@

&{}+\frac{2}{5}g_1^2g_1'^2\left [ 3Q_2^2+\frac{1}{3}Q_Q^2+\frac{16}{3}Q_u^2+\left ( 3Q_2+Q_Q-4Q_u\right )\Sigma_Q^Y\right ]\bigg \},\label{eq:USSMYu22BetaTwoLoop}\\
